---
title: "Pileup Plots for Khd1 analysis"
author: "Jing Qi Chong"
output: html_document
---

## Summary
Generates pileup plot for Khd1 CRAC data on specific regions of specific Khd1 target genes that are specified in `~/CRAC_analysis_2022/input_annotation/Khd1_TargetGeneNamesOnly.txt`. These show detailed profiles of read counts, including nucleotide-specific mutations and deletions, along selected transcripts.

This script relies on "pileup" files in tab-separated text format produced by `pyPileup` script while running the pipeline, put in the directory `~/CRAC_analysis_2022/khd1/khd1_SRR847751/pyPileup_analyses/`. So it must be run after the nf_CRACpipeline, read the repository `README.md` for details.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache= TRUE,
                      cache.path = "cache/pileup_Khd1-",
                      fig.path = "figure/pileup_Khd1-")
library(ggplot2)
library(readr)
library(tidyr)
library(dplyr)
library(forcats)
library(stringr)
library(cowplot)
theme_set(theme_cowplot(font_size = 11) + 
              theme(strip.background = element_blank()) 
          )
```

## Functions for data loading and plotting

```{r pileup_functions}
read_pileuptxt <- function(file,converturidine=FALSE) {
    pileupdf <- readr::read_tsv(file = file,
                    comment = "#",
                    col_types = "cifiii",
                    col_names = c("target","position","nucleotide",
                                  "hits","substitutions","deletions")
    )
    if (converturidine) {
        pileupdf <- mutate(pileupdf,
                           nucleotide = fct_recode(nucleotide, U = "T") )
    }
    pileupdf
}

scale_fill_pileup <- function(...) {
    scale_fill_manual(values = c(aligned = "grey50",
                                 substitutions = "black",
                                 deletions = "red2"),
                      ...)
}

plot_pileup_one <- function(pileupdf, 
                            include = "allonly",
                            nucseq = FALSE, 
                            nuctxtsize = 2.8) {
    p_pileup <- ggplot(data = pileupdf, aes(x=position)) +
        coord_cartesian(expand = FALSE, clip = "off")
    if (include == "allonly") {
        p_pileup <- p_pileup + 
            geom_col(aes(y = hits))
    } else if (include == "allsub") {
        p_pileup <- p_pileup + 
            geom_col(aes(y = hits,
                         fill = "aligned")) +
            geom_col(aes(y = substitutions,
                         fill = "substitutions")) 
    } else if (include == "alldel") {
        p_pileup <- p_pileup + 
            geom_col(aes(y = hits,
                         fill = "aligned")) + 
            geom_col(aes(y = deletions,
                         fill = "deletions"))
    } else if (include == "allsubdel") {
        # this is a hack that prints substitutions behind deletions
        # so that total height is added, it looks "stacked".
        # There would be a better way of doing it with reshaping and
        # position = "stack".
        p_pileup <- p_pileup + 
            geom_col(aes(y = hits,
                         fill = "aligned"))  +
            geom_col(aes(y = substitutions + deletions,
                         fill = "substitutions")) + 
            geom_col(aes(y = deletions,
                         fill = "deletions"))  + 
            scale_fill_pileup("Hit type")
    }
    
    if (nucseq) {
        p_pileup <- p_pileup + 
            geom_text(aes(label = nucleotide),
                      y = 0, family = 'Courier', vjust = 1.1,
                      size = nuctxtsize) +
            theme(axis.title.x =  element_blank(),
                  axis.line.x = element_blank(),
                  axis.text.x = element_blank(),
                  axis.ticks.x = element_blank()) +
            expand_limits(y = - max(pileupdf$hits / 16 ) )
    }
    p_pileup
}

plot_pileuphits_line <- function(pileupdf) {
    ggplot(data = pileupdf, 
           aes(x=position, y = hits)) +
        coord_cartesian(expand = FALSE) + 
        geom_line(aes(y = hits))
}

plot_nucseq <- function(nucdf) {
    # fix the width
    ggplot(data = nucdf, aes(x = position, label = nucleotide) ) +
        geom_text(y=0,family='Courier') + 
        theme_nothing()
}

CNN_colour <- "#3fa6b0"
Cross_linking_colour <- "#ea1111"

geom_motifhighlightpt <- function(...) {
    geom_point(y=0, aes(x=mid), colour = CNN_colour, size = 3, shape = 17, vjust = 1,
               ...)
}

annotate_motifhighlightsegment <- function(mid,halfwidth=3.5) {
    annotate(y=0, yend = 0, x = mid-halfwidth, xend=mid+halfwidth,
             geom="segment", colour = CNN_colour, size = 2)
}

annotate_motifhighlight <- function(mid,...) {
    annotate(geom="text", size = 3, y = 0, x = mid, vjust = 0.9, fontface = "bold",
             ...)
}

annotate_CNN <- function(mid, ...) {
    annotate_motifhighlight(mid, label = "C", colour = CNN_colour, ...)
}

annotate_CS <- function(mid, ...) {
    annotate_motifhighlight(mid, label = "CS", colour = Cross_linking_colour, ...)
}

```

## Load khd1_SRR847751 sample for analysis
```{r load_data}
Khd1_SRR847751_df <- read_pileuptxt("~/CRAC_analysis_2022/khd1/khd1_SRR847751/pyPileup_analyses/SRR847751_trimmed_pileups.txt",converturidine = TRUE)
Khd1_SRR847751_filtered_4_df <- Khd1_SRR847751_df %>%
  mutate(percent_del = (deletions / hits) * 100) %>%
  filter( hits > 100, percent_del >= 4)
```

##Generate table containing the genes exceeding the threshold
```{r generate table}
Khd1_targetted_genes <- data.frame((unique(sort(Khd1_SRR847751_filtered_4_df$target))))
```

##ASH1
## Plot pileup along ASH1 transcript as a line
```{r plot_ASH1line,dependson = c("pileup_functions","load_data"),fig.height = 2,fig.width=4}
plot_pileuphits_line(filter(Khd1_SRR847751_df, target == "ASH1") )
```
## Plot pileup focus on peaks ASH1
```{r plot_ASH1focus,dependson = c("pileup_functions","load_data"),fig.height = 4,fig.width=6}
ASH1wide_df <- filter(Khd1_SRR847751_df, 
                         target == "ASH1", 
                         position >= 700, 
                         position <= 1000)

ASH1wide_plot <- plot_pileup_one(ASH1wide_df, include = "allsubdel") + 
    labs(title = "ASH1 , pileup plot") 
    #annotate_CS(filter(Khd1_SRR847751_filtered_5_df,target=="ASH1")$position)
ASH1wide_plot
##ggsave(plot = ASH1wide_plot, filename = "../khd1_SRR847751/figure_pileup/ASH1_CS_plot.pdf",
#       height = 3, width=6, device = "pdf")
```


## Plot pileup focus on ASH1 specific coding region

```{r plot_ASH1seq,dependson = c("pileup_functions","load_onetable"),fig.height = 3, fig.width=8}
ASH1seq_df <- filter(Khd1_SRR847751_df, 
                      target == "CCW12", 
                      position >= 0, 
                      position <= 120)
ASH1seq_plot <- plot_pileup_one(ASH1seq_df, 
                                   include = "allsubdel", 
                                   nucseq = TRUE, nuctxtsize = 2) + 
    labs(title = "CCW12 pileup plot")
ASH1seq_plot
ggsave(plot = ASH1seq_plot, filename = "../khd1_SRR847751/figure_pileup/CCW12_seqplot.png",
       height = 3, width=10, device = "png")
```  

## Plot pileups centred only on Khd1-enriched motifs

Plot pileup on same area surrounding the CNN motifs (-12, +2) to look for consistent location of crosslinking sites.

```{r CNN_positions_all,dependson = c("pileup_functions","load_onetable"),fig.height = 12, fig.width =12}
CNN_positions_all <- position_locate_all(Khd1_SRR847751_df) %>%
    as_tibble() %>%
    mutate(motifid = 1:nrow(.))

get_chunk <- function(sedf,datadf) {
    datadf[sedf$start : sedf$end,]
}

CNN_motif_pileups <- 
    CNN_positions_all[-1,] %>%
    mutate(start = start - 12, end=end + 2) %>%
    group_by(motifid) %>%
    do(get_chunk(., datadf = Khd1_SRR847751_df)) %>%
    mutate(motifidg = paste(target, motifid),
           maxhits = max(hits)) %>%
    filter(maxhits > 150)

CNN_motif_pileups %>%
    do(plot = plot_pileup_one(.,include = "allsubdel",
                              nucseq = TRUE,nuctxtsize = 2) +
        theme(legend.position = "none") + 
           labs(title = .$motifidg[1])) %>%
    .$plot %>%
    plot_grid(plotlist = .)
# CNN_motifs
```


## Find CNN motif positions separately for each transcript

```{r CNN_positions_bytx,dependson = c("pileup_functions","load_onetable"),fig.height = 8, fig.width = 8}
position_locate_all <- function(posdf,
                                pattern = "C[ACUG][ACUG]C[ACUG][ACUG]C[ACUG][ACUG]") {
    posdf$nucleotide %>%
        as.character() %>%
        paste(collapse = "") %>%
        stringr::str_locate_all(pattern) %>%
        .[[1]]
}

CNN_positions_bytx <- Khd1_SRR847751_df %>%
    group_by(target) %>%
    do(., position_locate_all(.) %>% as_tibble() ) %>%
    mutate(mid = start + 3.5)
CNN_positions_bytx

```
##Plot pileup near Khd1 cross-linked sites
```{r plot_ASH1_reproducibility}
#ASH1
ASH1_widerep_df <- filter(Khd1_SRR847751_df, 
                         target == "ASH1",
                         position >= 820, 
                         position <= 880 )

ASH1_widerep_plot <- 
    plot_pileup_one(ASH1_widerep_df, include = "allsubdel",nucseq = TRUE, nuctxtsize = 3) + 
    labs(title = "ASH1", x = "position from TSS") +
  xlim(min(ASH1_widerep_df$position) - 1,max(ASH1_widerep_df$position) + 1) +
  theme(plot.margin = margin(t=10,b = 25)) 
  #  annotate_CNN(filter(CNN_positions_bytx,target=="ASH1")$mid) 

#SRL1
SRL1_widerep_df <- filter(Khd1_SRR847751_df, 
                         target == "SRL1",
                         position >= 550, 
                         position <= 625 )

SRL1_widerep_plot <- 
    plot_pileup_one(SRL1_widerep_df, include = "allsubdel",nucseq = TRUE, nuctxtsize = 3) + 
    labs(title = "SRL1", x = "position from TSS") +
  xlim(min(SRL1_widerep_df$position) - 1,max(SRL1_widerep_df$position) + 1) +
  theme(plot.margin = margin(t=25, b= 25))
  #  annotate_CNN(filter(CNN_positions_bytx,target=="SRL1")$mid) 

#CCW12 MOTIF 1
CCW12_1_widerep_df <- filter(Khd1_SRR847751_df, 
                         target == "CCW12",
                         position >= 210, 
                         position <= 270 )

CCW12_1_widerep_plot <- 
    plot_pileup_one(CCW12_1_widerep_df, include = "allsubdel",nucseq = TRUE, nuctxtsize = 3) + 
    labs(title = "CCW12_1", x = "position from TSS") +
  xlim(min(CCW12_1_widerep_df$position) - 1,max(CCW12_1_widerep_df$position) + 1) +
  theme(plot.margin = margin(t= 25, b= 25))
  #  annotate_CNN(filter(CNN_positions_bytx,target=="CCW12")$mid) 

#CCW12 MOTIF 2
CCW12_2_widerep_df <- filter(Khd1_SRR847751_df, 
                         target == "CCW12",
                         position >= 400, 
                         position <= 460 )

CCW12_2_widerep_plot <- 
    plot_pileup_one(CCW12_2_widerep_df, include = "allsubdel",nucseq = TRUE, nuctxtsize = 3) + 
    labs(title = "CCW12_2", x = "position from TSS") +
  xlim(min(CCW12_2_widerep_df$position) - 1,max(CCW12_2_widerep_df$position) + 1) +
  theme(plot.margin = margin(t= 25))
  #  annotate_CNN(filter(CNN_positions_bytx,target=="CCW12")$mid) 

pileup_plot_allgenes <- 
  plot_grid(ASH1_widerep_plot,SRL1_widerep_plot,CCW12_1_widerep_plot,CCW12_2_widerep_plot, ncol = 1) 


pileup_plot_allgenes

ggsave(plot = pileup_plot_allgenes, 
        filename = "../khd1_SRR847751/figure_pileup/Composite_CS_CNN_pileupplot.png",
        height = 8, width = 9, device = "png")

```

##Plot pileup near Khd1 cross-linked sites
```{r plot_ASH1_reproducibility}
HSP150_widerep_df <- filter(Khd1_SRR847751_df,target== "HSP150") 
       

HSP150_widerep_plot <- 
    plot_pileup_one(ASH1_widerep_df, include = "allsubdel") + 
    labs(title = "HSP150 peak", x = "position from TSS")  +
    xlim(min(ASH1_widerep_df$position) - 1,max(ASH1_widerep_df$position) + 1) +
    annotate_CNN(filter(CNN_positions_bytx,target=="HSP150")$mid) 
HSP150_widerep_plot

```

## Session Info

Package versions, etc., that ran this file.

```{r session_info,results="show",message=TRUE}
sessionInfo()
```
